Title:run_analysis.R
========================================================
The purpose of this project is to demonstrate your ability to collect, work with, and clean a data set. The goal is to prepare tidy data that can be used for later analysis. You will be graded by your peers on a series of yes/no questions related to the project. You will be required to submit: 1) a tidy data set as described below, 2) a link to a Github repository with your script for performing the analysis, and 3) a code book that describes the variables, the data, and any transformations or work that you performed to clean up the data called CodeBook.md. You should also include a README.md in the repo with your scripts. This repo explains how all of the scripts work and how they are connected.  

One of the most exciting areas in all of data science right now is wearable computing - see for example this article . Companies like Fitbit, Nike, and Jawbone Up are racing to develop the most advanced algorithms to attract new users. The data linked to from the course website represent data collected from the accelerometers from the Samsung Galaxy S smartphone. A full description is available at the site where the data was obtained: 

http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones 

Here are the data for the project: 

https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip 

 You should create one R script called run_analysis.R that does the following. 
Merges the training and the test sets to create one data set.
Extracts only the measurements on the mean and standard deviation for each measurement. 
Uses descriptive activity names to name the activities in the data set
Appropriately labels the data set with descriptive activity names. 
Creates a second, independent tidy data set with the average of each variable for each activity and each subject. 

The first step is to load all the relevant files 
```{r}
##Loads the Test files
x_test <- read.table("test/X_test.txt")
y_test <-read.table("test/Y_test.txt")
subs_test<-read.table("test/subject_test.txt")
##Loads Training Files
x_train <- read.table("train/X_train.txt")
y_train <-read.table("train/Y_train.txt")
subs_train<-read.table("train/subject_train.txt")
##Loads shared files 
features <- read.table("features.txt")
activities <- read.table("activity_labels.txt")
```

Once we have the file we then combine them so we can start working with them
```{r}
#this will combine the files into the alls
X_all<- (rbind(x_test,x_train))
Y_all<- (rbind(y_test,y_train))
all_subs<-(rbind(subs_test,subs_train))
```

We then work a bit on fixing the column names for the data
```{r}
##users features to replace the column names
colnames(X_all)<-features$V2

#adds info from activity files to the Y_All data_frame
#Also adds a nice col name to the activity
Y_all_act<- merge(Y_all,activities)
colnames(Y_all_act)[2]<-"Activity"
#adds a nice col name for the subject file
colnames(all_subs)[1]<-"Subject"
```
Now will we be combining all the data we have been working on into one data frame with 

```{r}
#combiles all the data files then removed an 
##unneed col which has row number in it.
Fin<-cbind(all_subs,Y_all_act,X_all)
Fin<-Fin[,c(1,3:564)]
```

Now we filter downt he columns to get the ones with mean and STD in them.

```{r}
#creates col names for what we want to focus on i.e STD and means
stdcols<-grep("std",colnames(Fin))
meancols<-grep("mean",colnames(Fin))

#filters the data for onlyt he columns we want whilst keeping
#the first two columns
Fin_f<-Fin[,c(1:2,stdcols,meancols)]
```

Finally we update col names once again and then output to a file.

```{r}
tmp<-aggregate(Fin_f[,3:81], by=list((Fin_f$Subject),Fin_f$Activity), FUN=mean)
##Fixing Col names again
colnames(tmp)[1]<-"Subject"
colnames(tmp)[2]<-"Activity"

#write to a file
write.table(tmp, "tidy_data_set.txt")
```

